---
title: "Pythium_biomass"
author: "Fiama Guevara"
date: "2024-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#### Upload libraries
library(readxl)
library(ggplot2)
library("lme4")
library("car")
library ("emmeans")
library("tidyverse")
library("tidyr")
library(ggpubr)
library(ggthemes)
library(multcomp) #poshoc lmer
library(ggbreak) #scale breaks
library(patchwork) #multiplots
library(ggpattern)
```

# Importing data
```{r}
#Import from Excel option used
library(readxl)
Pathogens_data <- read_excel("Pathogens_data.xlsx", 
    sheet = "Pythium_Biomass")

View(Pathogens_data)
```

# Mean biomass all groups
```{r}
#Mean by groups
meansBiomass <- Pathogens_data %>%
  group_by(Group, Treatment, sample) %>%
  summarise_at(vars(c(Fresh_Biomass, Dry_Biomass)), list(name = mean, sd))

View(meansBiomass)

```

# Analyses BIOMASS samples
## Roots fresh weight
```{r}
#Sub-setting only Fresh weight roots
data.Roots <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots")

View(data.Roots)

#Mix models - experiment run as random variable
model.FWR <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run), data = data.Roots)

Anova(model.FWR) #Significant effect treatment p.value= 0.01896 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR, which=1:6) 
res <- resid(model.FWR)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.3115

# Poshoc
emmeans(model.FWR, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between Control - Inoculated p-value=0.0218*

emmeans(model.FWR, list(pairwise ~ Treatment), adjust = "bonferroni")

```

## Roots dry weight
```{r}

#Mix models - experiment run as random variable
model.DWR <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run), data = data.Roots)

Anova(model.DWR) #NO Significant effect treatment p.value= 0.4417

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWR, which=1:6) 
resD <- resid(model.DWR)

par(mfrow = c(1, 2)) # combine plots
hist(resD, breaks = 40)

# QQ-plot
qqPlot(resD,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resD) #NOT Normal p-value= 0.03764

# Poshoc
emmeans(model.DWR, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between Control - Inoculated p-value=0.4443

compare_means(Dry_Biomass ~ Treatment, data = data.Roots, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format=0.39 and by run2=0.311 - run3=0.018*

```

##Leaves fresh weight
```{r}
#Sub-setting only Fresh weight roots
data.Leaves <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves")

View(data.Leaves)


#Mix models - experiment run as random variable
model.FWL <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run), data = data.Leaves)

Anova(model.FWL) #Significant effect treatment p.value= 0.007849 **

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWL, which=1:6) 
resL <- resid(model.FWL)

par(mfrow = c(1, 2)) # combine plots
hist(resL, breaks = 40)

# QQ-plot
qqPlot(resL,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resL) #ALMOST Normal p-value= 0.04973 (one outlier)

# Poshoc
emmeans(model.FWL, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between Control - Inoculated p-value=0.0097**

compare_means(Fresh_Biomass ~ Treatment, data = data.Leaves, method = "t.test", group.by = "Experiment_run")  # Without grouping p.format=0.01* and by run2=0.4429 - run3=0.0042*

compare_means(Fresh_Biomass ~ Treatment, data = data.Leaves, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format=0.014* and by run2=0.376 - run3=0.012*

```
##Leaves dry weight
```{r}

#Mix models - experiment run as random variable
model.DWL <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run), data = data.Leaves)

Anova(model.DWL) #Significant effect treatment p.value= 0.04536 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWL, which=1:6) 
resDL <- resid(model.DWL)

par(mfrow = c(1, 2)) # combine plots
hist(resDL, breaks = 40)

# QQ-plot
qqPlot(resDL,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resDL) #NOT Normal p-value= 0.002365

# Poshoc
emmeans(model.DWL, list(pairwise ~ Treatment))
#Significant diff between Control - Inoculated p-value=0.0493

#Non-parametric test
compare_means(Dry_Biomass ~ Treatment, data = data.Leaves, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format=0.022* and by run2=0.547 - run3=0.0056*


# Conduct pairwise comparisons with Wilcoxon rank-sum test
pairwise_result2 <- pairwise.wilcox.test(data.Leaves$Fresh_Biomass, data.Leaves$Treatment, p.adjust.method = "bonferroni")
print(pairwise_result2) #p-value=0.014

```

## Combined boxplot Biomass
```{r}
#Sub-setting only BIOMASS
data.Biomass <- Pathogens_data %>% filter(Group == "BIOMASS")

# Final plot Fresh Biomass
combined.plot <- ggplot(data.Biomass, aes(x = sample, y = Fresh_Biomass, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Pythium inoculation", values = c("#1F78B4", "orchid")) +
  labs(x = "\nSample type", y = "\nFresh biomass (g)\n") +
      theme_bw() +
  scale_y_cut(breaks=37, scales=c(0, 5))


# Save plot
ggsave("Fresh_Biomass_Pythium.png", combined.plot, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```

```{r}

# Final plot Dry Biomass
combined.plotD <- ggplot(data.Biomass, aes(x = sample, y = Dry_Biomass, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Pythium inoculation", values = c("#1F78B4", "orchid")) +
  labs(x = "\nSample type", y = "\nDry biomass (g)\n") +
      theme_bw() + 
        scale_y_cut(breaks=3, scales=c(0, 0.5)) 


# Save plot
ggsave("Dry_Biomass_Pythium.png", combined.plotD, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```

# Analyses DNA samples
```{r}
meansBiomass2 <- Pathogens_data %>% filter(Group == "DNA") %>%
  group_by(Visual_dis_severity, sample) %>%
  summarise_at(vars(c(Fresh_Biomass, Dry_Biomass)), list(name = mean, sd))

View(meansBiomass2)
```

## Fresh Roots
```{r}
##### Fresh weight roots #####
data.DNAr <- Pathogens_data %>% filter(Group == "DNA", sample == "Roots")

View(data.DNAr)

#Mix model
model.FWR23 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Experiment_run), data = data.DNAr)

Anova(model.FWR23) #Significant effect of Disease severity p-value=1.239e-07 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR23, which=1:6) 
res <- resid(model.FWR23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal  p-value = 0.001273

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.FWR23, "Visual_dis_severity", )
pairs(mSev) #significant effect disease severity

multcomp::cld(object = mSev, Letters = letters)
###Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  16.2 3.18 1.13   -14.58     46.9  a    
 #Control               21.2 3.11 1.04   -14.70     57.1   b   
 #Low                   23.0 3.18 1.13    -7.74     53.8   b

# Poshoc
emmeans(model.FWR23, list(pairwise ~ Visual_dis_severity), adjust= "bonferroni")
#Significant diff between Control - high p-value=0.0001

#Non-parametric test
compare_means(Fresh_Biomass ~ Visual_dis_severity, data = data.DNAr, method = "kruskal.test") ## Without grouping p.format=

compare_means(Fresh_Biomass ~ Visual_dis_severity, data = data.DNAr, method = "kruskal.test", group.by = "Experiment_run")  # by run2=0.0017** - run3=0.0252*

# Conduct pairwise comparisons with Wilcoxon rank-sum test
pairwise_result <- pairwise.wilcox.test(data.DNAr$Fresh_Biomass, data.DNAr$Visual_dis_severity, p.adjust.method = "bonferroni")
print(pairwise_result)


```



## Fresh Leaves
```{r}
##### Fresh weight roots #####
data.DNAl <- Pathogens_data %>% filter(Group == "DNA", sample == "Leaves")

model.FWLt <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run), data = data.DNAl)
Anova(model.FWLt) #Sigificant effect treatment p-value=0.01256 *

model.FWL23 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Experiment_run), data = data.DNAl)

Anova(model.FWL23) #Significant effect of Disease severity p-value=< 2.2e-16 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWL23, which=1:6) 
res <- resid(model.FWL23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value = 0.7079

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.FWL23, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low and high-control
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  77.5 13.0 1.49    -1.12      156  a    
 #Control              136.3 12.2 1.15    22.65      250   b   
 #Low                  150.6 13.0 1.49    71.95      229   b 

```


## Combined boxplot DNA FWR&FWL
```{r}
#Sub-setting only BIOMASS
data.DNA <- Pathogens_data %>% filter(Group == "DNA")

### Dataset imported from Excel 
#To reorder the factor  
data.DNA$Visual_dis_severity <- ordered(data.DNA$Visual_dis_severity, levels = c("Low", "Control", "High"))


# Final plot 
combined.plot.DNA <- ggplot(data.DNA, aes(x = Visual_dis_severity, y = Fresh_Biomass, fill = Visual_dis_severity)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Disease severity", values = c("palegreen1", "#1F78B4", "hotpink1")) +
      labs(x= "\nDisease severity", y = "\nFresh biomass (g)\n", title= "Samples for DNA extraction") +
  facet_grid(~sample) + theme_classic() +
    theme(legend.position = "none") + 
      scale_y_cut(breaks=37, scales=c(0, 5))

#outlier.shape = NA

# Save plot
ggsave("FWR_FWL.png", combined.plot.DNA, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```


## Analysis Root length run 2&3 (NOT UPDTAED)
```{r}
library(readxl)
Pathogens_data <- read_excel("Pathogens_data.xlsx", 
    sheet = "Pythium_all")
View(Pathogens_data)
```

```{r}

##### Fresh weight roots #####
data.RL <- Pathogens_data %>% filter(Group == "DNA", Experiment_run != "Run1")

model.RL23 <- lmer(Root_length ~ Visual_dis_severity + (1|Experiment_run), data = data.RL)

Anova(model.RL23) #Significant effect of Disease severity p-value=0.02457 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.RL23, which=1:6) 
res <- resid(model.RL23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal   p-value = 0.7118

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.RL23, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low
multcomp::cld(object = mSev, Letters = letters)

emmeans(model.RL23, list(pairwise ~ Visual_dis_severity), adjust = "tukey")

```
### Box plot run2&3
```{r}
#To reorder the factor  
data.RL$Visual_dis_severity <- ordered(data.RL$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
ColorsBP <- c("palegreen1", "#1F78B4", "hotpink1")

RL23.plot <- ggboxplot(data.RL, x = "Visual_dis_severity", y = "Root_length",
          fill = "Visual_dis_severity", palette = ColorsBP, outlier.shape = NA) +
      labs(x = "\nDisease severity", y = "Root length (cm)") +
        ylim(25, 100) +
  theme(legend.position = "none") 

# Save plot
ggsave("RL_DNAsamples.png", RL23.plot, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```


#Analysis Run 1 only
## Roots FW run 1
```{r}
##### Fresh weight roots #####
DNA_Rsamples1 <- Pathogens_data %>% filter(Group == "DNA", Experiment_run == "Run1")

model.FWR1 <- lmer(Roots_fresh_weight ~ Visual_dis_severity + pH + EC + (1|Pond), data = DNA_Rsamples1)

Anova(model.FWR1) #No Significant effects

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR1, which=1:6) 
res <- resid(model.FWR1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Not normal  

#Perform the test
compare_means(Roots_fresh_weight ~ Visual_dis_severity, data = DNA_Rsamples1, method = "kruskal.test")  #group.by = "Run"

```
### Box plot run1
```{r}
#To reorder the factor  
DNA_Rsamples1$Visual_dis_severity <- ordered(DNA_Rsamples1$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
FWR1.plot <- ggboxplot(DNA_Rsamples1, x = "Visual_dis_severity", y = "Roots_fresh_weight",
          fill = "Visual_dis_severity", palette = "lancet") +
      labs(x = "\nDisease severity", y = "Roots fresh weight (g)") +
        ylim(15, 50) +
  theme(legend.position = "none")


# Use only p.format as label. Remove method name.
FWR1.plot + stat_compare_means(label = "p.format", method = "kruskal.test", label.y = 48) +
  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Control", label.y = 42) 


```

## Roots FW run 2
```{r}
##### Fresh weight roots #####
DNA_Rsamples2 <- Pathogens_data %>% filter(Group == "DNA", Experiment_run == "Run2")

model.FWR2 <- lmer(Roots_fresh_weight ~ Visual_dis_severity + Temperature + pH + EC + (1|Pond), data = DNA_Rsamples2)

Anova(model.FWR2) #Significant effect of Disease

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR2, which=1:6) 
res <- resid(model.FWR2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Not normal  

#Perform the test
compare_means(Roots_fresh_weight ~ Visual_dis_severity, data = DNA_Rsamples2, method = "kruskal.test")  #group.by = "Run"

```
### Box plot run2
```{r}
#To reorder the factor  
DNA_Rsamples2$Visual_dis_severity <- ordered(DNA_Rsamples2$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
FWR2.plot <- ggboxplot(DNA_Rsamples2, x = "Visual_dis_severity", y = "Roots_fresh_weight",
          fill = "Visual_dis_severity", palette = "lancet") +
      labs(x = "\nDisease severity", y = "Roots fresh weight (g)") +
        ylim(10, 50) +
  theme(legend.position = "none")


# Use only p.format as label. Remove method name.
FWR2.plot + stat_compare_means(label = "p.format", method = "kruskal.test", label.y = 48) +
  stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = "Control", label.y = 42) 


```
## Roots FW run 3
```{r}
##### Fresh weight roots #####
DNA_Rsamples3 <- Pathogens_data %>% filter(Group == "DNA", Experiment_run == "Run3")

model.FWR3 <- lmer(Roots_fresh_weight ~ Visual_dis_severity + Temperature + pH + EC + (1|Pond), data = DNA_Rsamples3)

Anova(model.FWR3) #Significant effect of Disease

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR3, which=1:6) 
res <- resid(model.FWR3)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal  

#Perform the test
# Poshoc
emmeans(model.FWR3, "Visual_dis_severity")


compare_means(Roots_fresh_weight ~ Visual_dis_severity, data = DNA_Rsamples3, method = "t.test")  #group.by = "Run"

```

### Box plot run3
```{r}
#To reorder the factor  
DNA_Rsamples3$Visual_dis_severity <- ordered(DNA_Rsamples3$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
FWR3.plot <- ggboxplot(DNA_Rsamples3, x = "Visual_dis_severity", y = "Roots_fresh_weight",
          fill = "Visual_dis_severity", palette = "lancet", outlier.shape = NA) +
      labs(x = "\nDisease severity", y = "Roots fresh weight (g)") +
        ylim(5, 30) +
  theme(legend.position = "none") 


```

```{r}
# Get all plots in one figure
final.FWRs_DNA <- ggarrange(FWR1.plot, FWR2.plot, FWR3.plot,
                ncol = 3,  align = "hv")

# Save plot
ggsave("FWR_DNArun2.png", final.FWRs_DNA, 
  width = 12,
  height = 11,
  units = "cm",
  dpi = 600)
```

#qPCR results
```{r}
#Import data
library(readxl)

qPCR_data <- read_excel("C:/Users/fiama/OneDrive - The Ohio State University/Hydroponics_Project/Greenhouse_experiments/pathogens_experiments/Pythium/qPCR/qPCR_Pythium_experiments_final.xlsx", 
    sheet = "Sheet1")

View(qPCR_data)
```

```{r}
#Change format numeric values
library(gt)
qPCR_data |>
  fmt_scientific(
    columns = Starting_Quantity,
    exp_style = "E")
```


## Starting qualtity
```{r}
#Test model 
model.qPCR <- lm(Starting_Quantity ~ Treatment * Exp_run * Replicate, data = qPCR_data) 

Anova(model.qPCR) #Significant effect interaction Treatment:Exp_run p.value= 0.008430 **
```


```{r}
##Analysis by run
#RUN 1
data_R1 <- qPCR_data %>% filter(Exp_run == "Run1")

#Model
model.qPCRr1 <- lmer(Starting_Quantity ~ Treatment * (1|Replicate), data = data_R1) 
Anova(model.qPCRr1) #Sig effect treatment p-value=0.0000001938 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.qPCRr1, which=1:6) 
res <- resid(model.qPCRr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.000004838

# Poshoc
emmeans(model.qPCRr1, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between all

emmeans(model.FWR, list(pairwise ~ Treatment), adjust = "bonferroni")

```


```{r}
##Analysis by run
#RUN 2
data_R2 <- qPCR_data %>% filter(Exp_run == "Run2")

#Model
model.qPCRr2 <- lmer(Starting_Quantity ~ Treatment * (1|Replicate), data = data_R2) 
Anova(model.qPCRr2) #Sig effect treatment p-value=0.0481 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.qPCRr2, which=1:6) 
res <- resid(model.qPCRr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=5.938e-14

# Poshoc
emmeans(model.qPCRr2, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between any


```

##Plot
```{r}
#Label x- axis scale
library(scales)
scientific <- function(x){
    ifelse(x==0, "0", parse(text=gsub("[+]", "", gsub("e", " %*% 10^", scientific_format()(x)))))
}


plotR1 <- ggplot(qPCR_data, aes(x = Exp_run, y = Starting_Quantity, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Disease severity", values = c("#1F78B4", "orchid", "green4")) +
  labs(x = "\nExperimental run", y = "\nStarting quantity (ng/uL)\n") +
      theme_bw() +
  scale_y_continuous(label=scientific)


# Save plot
ggsave("qPCR_plot_StartinQuantity.png", plotR1, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```

## Cq
A lower Cq value indicates a higher amount of genetic material in a sample. 
```{r}
#Test model 
model.qPCR2 <- lm(Cq ~ Treatment * Exp_run * Replicate, data = qPCR_data) 

Anova(model.qPCR2) #Significant effect interaction Treatment:Exp_run p.value= 0.0000067017260
```


```{r}
##Analysis by run
#RUN 1
data_R1 <- qPCR_data %>% filter(Exp_run == "Run1")

#Model
model.qPCRr1c <- lmer(Cq ~ Treatment * (1|Replicate), data = data_R1) 
Anova(model.qPCRr1c) #Sig effect treatment p-value=0.000000006717 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.qPCRr1c, which=1:6) 
res <- resid(model.qPCRr1c)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.00000052

# Poshoc
emmeans(model.qPCRr1c, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between control-low and control-high


```

```{r}
##Analysis by run
#RUN 2
data_R2 <- qPCR_data %>% filter(Exp_run == "Run1")

#Model
model.qPCRr2c <- lmer(Cq ~ Treatment * (1|Replicate), data = data_R2) 
Anova(model.qPCRr2c) #NO Sig effect treatment p-value=0.125

#Residuals plots
par(mfrow = c(2, 3))
plot(model.qPCRr2c, which=1:6) 
res <- resid(model.qPCRr2c)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.2376

# Poshoc
emmeans(model.qPCRr2c, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between any


```

##Plot Cq
```{r}

plotR1c <- ggplot(qPCR_data, aes(x = Exp_run, y = Cq, fill = Treatment)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Disease severity", values = c("#1F78B4", "orchid", "green4")) +
  labs(x = "\nExperimental run", y = "\nQuantification cycle (Cq)\n") +
      theme_bw() 


# Save plot
ggsave("qPCR_plot_Cq.png", plotR1c, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```


