---
title: "Phytium_biomass_V2"
author: "Fiama Guevara"
date: "2026-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#### Upload libraries
library(readxl)
library(ggplot2)
library("lme4")
library("car")
library ("emmeans")
library("tidyverse")
library("tidyr")
library(ggpubr)
library(ggthemes)
library(multcomp) #poshoc lmer
library(ggbreak) #scale breaks
library(patchwork) #multiplots
library(ggpattern)
```

# REVIEWED ANALYSIS (04/10/2026) - for Phytofrontiers reviewers comments
## Importing data
```{r}
#Import from Excel option used
library(readxl)
Pathogens_data <- read_excel("Pythium_Biomass_dataset.xlsx")

View(Pathogens_data)

```

# Analyses DNA samples
```{r}
meansBiomass2 <- Pathogens_data %>% filter(Group == "DNA") %>%
  group_by(Visual_dis_severity, sample) %>%
  summarise_at(vars(c(Fresh_Biomass, Dry_Biomass, Root_length)), list(name = mean, sd))

View(meansBiomass2)
```

```{r}
# ANOVA to check effect of all variables
data.DNAr <- Pathogens_data %>% filter(Group == "DNA", sample == "Roots")

#Simple model
model.FWR <- lm(Fresh_Biomass ~ Visual_dis_severity *Treatment * Experiment_run *Pond, data = data.DNAr)

Anova(model.FWR) 

#Result
#                                                     Pr(>F)    
#Visual_dis_severity                               1.199e-08 ***
#Treatment                                                      
#Experiment_run                                    1.241e-11 ***
#Pond                                              2.192e-06 ***
#Visual_dis_severity:Treatment                                  
#Visual_dis_severity:Experiment_run                  0.01379 *  
#Treatment:Experiment_run                                       
#Visual_dis_severity:Pond                            0.53070    
#Treatment:Pond                                                 
#Experiment_run:Pond                                 0.66157    
#Visual_dis_severity:Treatment:Experiment_run                   
#Visual_dis_severity:Treatment:Pond                             
#Visual_dis_severity:Experiment_run:Pond             0.81583    
#Treatment:Experiment_run:Pond                                  
#Visual_dis_severity:Treatment:Experiment_run:Pond              
#Residuals                                                      
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

```


## Fresh Roots
```{r}
##### Fresh weight roots #####
data.DNAr <- Pathogens_data %>% filter(Group == "DNA", sample == "Roots")

View(data.DNAr)

#Mix model
model.FWR23 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Experiment_run/Pond), data = data.DNAr)

Anova(model.FWR23) #Result

# Summary with p-values
summary(model.FWR23)

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR23, which=1:6) 
res <- resid(model.FWR23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal  p-value = 0.007132

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.FWR23, "Visual_dis_severity", )
pairs(mSev) #significant effect disease severity

multcomp::cld(object = mSev, Letters = letters)
###Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  16.2 3.23 1.20   -11.66     44.0  a    
 #Control               21.2 3.18 1.14    -9.27     51.6   b   
 #Low                   23.0 3.23 1.20    -4.83     50.9   b  

# Poshoc
emmeans(model.FWR23, list(pairwise ~ Visual_dis_severity), adjust= "bonferroni")
#Significant diff between Control - high p-value=0.0383 and HIgh-Low pval= <.0001

#Non-parametric test
compare_means(Fresh_Biomass ~ Visual_dis_severity, data = data.DNAr, method = "kruskal.test") ## Without grouping p.format= 0.00047

compare_means(Fresh_Biomass ~ Visual_dis_severity, data = data.DNAr, method = "kruskal.test", group.by = "Experiment_run")  # by run2=0.0017** - run3=0.0252*

# Conduct pairwise comparisons with Wilcoxon rank-sum test
pairwise_result <- pairwise.wilcox.test(data.DNAr$Fresh_Biomass, data.DNAr$Visual_dis_severity, p.adjust.method = "bonferroni")
print(pairwise_result)

#      Control High   
#High 0.00228 -      
#Low  1.00000 0.00085

```

```{r}
# ANOVA to check effect of all variables
data.DNAl <- Pathogens_data %>% filter(Group == "DNA", sample == "Leaves")

#Simple model
model.FWL <- lm(Fresh_Biomass ~ Visual_dis_severity * Treatment * Experiment_run *Pond, data = data.DNAl)

Anova(model.FWL)

#Result
#Response: Fresh_Biomass
#                                                  Sum Sq Df F value    Pr(>F)    
#Visual_dis_severity                                48063  1 97.8478 1.004e-13 ***
#Treatment                                                 0                      
#Experiment_run                                      9959  1 20.2748 3.621e-05 ***
#Pond                                                6476  4  3.2960   0.01729 *  
#Visual_dis_severity:Treatment                             0                      
#Visual_dis_severity:Experiment_run                 10220  1 20.8057 2.963e-05 ***
#Treatment:Experiment_run                                  0                      
#Visual_dis_severity:Pond                            5310  2  5.4048   0.00725 ** 
#Treatment:Pond                                            0                      
#Experiment_run:Pond                                  942  4  0.4794   0.75069    
#Visual_dis_severity:Treatment:Experiment_run              0                      
#Visual_dis_severity:Treatment:Pond                        0                      
#Visual_dis_severity:Experiment_run:Pond              738  2  0.7514   0.47659    
#Treatment:Experiment_run:Pond                             0                      
#Visual_dis_severity:Treatment:Experiment_run:Pond         0                      
#Residuals                                          26525 54                      
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

```


## Fresh Leaves
```{r}
##### Fresh weight roots #####
data.DNAl <- Pathogens_data %>% filter(Group == "DNA", sample == "Leaves")

model.FWLt <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run/Pond), data = data.DNAl)

Anova(model.FWLt) #Sigificant effect treatment p-value=0.01256 *

model.FWL23 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Experiment_run/Pond), data = data.DNAl)

Anova(model.FWL23) #Significant effect of Disease severity p-value=< 2.2e-16 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWL23, which=1:6) 
res <- resid(model.FWL23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value = 0.7664

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.FWL23, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low and high-control
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  77.5 13.0 1.51    -0.16      155  a    
 #Control              136.3 12.2 1.17    26.01      247   b   
 #Low                  150.6 13.0 1.51    72.92      228   b                                         

```

## Combined boxplot DNA FWR&FWL
```{r}
#Sub-setting only BIOMASS
data.DNA <- Pathogens_data %>% filter(Group == "DNA")

### Dataset imported from Excel 
#To reorder the factor  
data.DNA$Visual_dis_severity <- ordered(data.DNA$Visual_dis_severity, levels = c("Low", "Control", "High"))


# Final plot 
combined.plot.DNA <- ggplot(data.DNA, aes(x = Visual_dis_severity, y = Fresh_Biomass, fill = Visual_dis_severity)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Disease severity", values = c("palegreen1", "#1F78B4", "hotpink1")) +
      labs(x= "\nDisease severity", y = "\nFresh biomass (g)\n", title= "Samples for DNA extraction") +
  facet_grid(~sample) + theme_classic() +
    theme(legend.position = "none") + 
      scale_y_cut(breaks=41, scales=c(0, 5))

#outlier.shape = NA

# Save plot
ggsave("FWR_FWL_byRun.png", combined.plot.DNA, 
  width = 15,
  height = 12,
  units = "cm",
  dpi = 600)

```

# Root length (combined run 2&3) 
```{r}
library(readxl)
Pathogens_data2 <- read_excel("Pathogens_RECORD_exps.xlsx",
    sheet = "Pythium_Biomass_dataset")
View(Pathogens_data2)
```

```{r}

##### Root length #####
data.RL <- Pathogens_data2 %>% filter(Group == "DNA", Experiment_run != "Run1")

model.RL23 <- lmer(Root_length ~ Visual_dis_severity + (1|Experiment_run/Pond), data = data.RL)

Anova(model.RL23) #Significant effect of Disease severity p-value=0.01645 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.RL23, which=1:6) 
res <- resid(model.RL23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal   p-value = 0.6147

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.RL23, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  57.0 4.11 2.40     41.9     72.2  a    
 #Control               60.1 3.74 1.64     40.1     80.0  ab   
 #Low                   66.7 4.11 2.40     51.5     81.8   b
 
emmeans(model.RL23, list(pairwise ~ Visual_dis_severity), adjust = "tukey")

#$`pairwise differences of Visual_dis_severity`
 #              estimate   SE   df t.ratio p.value
 #Control - High     3.01 3.92 13.6   0.768  0.7279
 #Control - Low     -6.64 3.92 13.6  -1.695  0.2424
 #High - Low        -9.65 3.42 59.0  -2.820  0.0177 **

```
## Box plot
```{r}
#To reorder the factor  
data.RL$Visual_dis_severity <- ordered(data.RL$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
ColorsBP <- c("palegreen1", "#1F78B4", "hotpink1")

RL23.plot <- ggboxplot(data.RL, x = "Visual_dis_severity", y = "Root_length",
          fill = "Visual_dis_severity", palette = ColorsBP, outlier.shape = NA) +
      labs(x = "\nDisease severity", y = "Root length (cm)") +
        ylim(25, 100) +
  theme(legend.position = "none") 

# Save plot
ggsave("RL_DNAsamples.png", RL23.plot, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```



# DNA By RUN
## Roots fresh weight
### Run 1 (former run2)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsr1 <- data.DNAr <- Pathogens_data %>% filter(Group == "DNA", sample == "Roots", Experiment_run == "Run2")

View(data.Rootsr1)

#Mix models - experiment run as random variable
model.FWRr1 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Pond), data = data.Rootsr1)

Anova(model.FWRr1) #Significant effect p.value= 3.495e-10 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWRr1, which=1:6) 
res <- resid(model.FWRr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.002805

# Poshoc
emmeans(model.FWRr1, list(pairwise ~ Visual_dis_severity), adjust = "tukey")
#Significant diff between High - Low p-val=<.0001

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Control - High     6.33 2.41  4.81   2.625  0.1035
 #Control - Low     -3.10 2.41  4.81  -1.286  0.4625
 #High - Low        -9.43 1.44 29.00  -6.562  <.0001

emmeans(model.FWRr1, list(pairwise ~ Visual_dis_severity), adjust = "bonferroni")

# Pairwise comparisons
mSev <- emmeans(model.FWRr1, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low only
multcomp::cld(object = mSev, Letters = letters)


```

### Run 2 (former run3)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsr2 <- data.DNAr <- Pathogens_data %>% filter(Group == "DNA", sample == "Roots", Experiment_run == "Run3")

View(data.Rootsr2)

#Mix models - experiment run as random variable
model.FWRr2 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Pond), data = data.Rootsr2)

Anova(model.FWRr2) #Significant effect p.value= 0.007503 **

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWRr2, which=1:6) 
res <- resid(model.FWRr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.1749 

# Poshoc
emmeans(model.FWRr2, list(pairwise ~ Visual_dis_severity), adjust = "tukey")
#Significant diff between High - Low p-val=0.0126

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Control - High    3.682 2.60  4.63   1.419  0.4051
 #Control - Low    -0.561 2.60  4.63  -0.216  0.9747
 #High - Low       -4.243 1.38 29.00  -3.065  0.0126*

emmeans(model.FWRr2, list(pairwise ~ Visual_dis_severity), adjust = "bonferroni")

# Pairwise comparisons
mSev <- emmeans(model.FWRr2, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low only
multcomp::cld(object = mSev, Letters = letters)


```

## Leaves fresh weight
### Run 1 (former run2)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesr1 <- Pathogens_data %>% filter(Group == "DNA", sample == "Leaves", Experiment_run == "Run2")

View(data.Leavesr1)

#Mix models - experiment run as random variable
model.FWLr1 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Pond), data = data.Leavesr1)

Anova(model.FWLr1) #Significant effect p.value= < 2.2e-16 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWLr1, which=1:6) 
res <- resid(model.FWLr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.01798

# Poshoc
emmeans(model.FWLr1, list(pairwise ~ Visual_dis_severity), adjust = "tukey")
#Significant diff between High - Low p-val=<.0001 

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Control - High     72.8 11.9  6.35   6.130  0.0017
 #Control - Low     -34.0 11.9  6.35  -2.862  0.0613
 #High - Low       -106.8 10.9 29.00  -9.809  <.0001

emmeans(model.FWLr1, list(pairwise ~ Visual_dis_severity), adjust = "bonferroni")

# Pairwise comparisons
mSev <- emmeans(model.FWLr1, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low and high-control
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  73.8 9.24 9.04       53     94.7  a    
 #Control              146.6 7.46 4.00      126    167.3   b   
 #Low                  180.6 9.24 9.04      160    201.5   b

```

### Run 2 (former run3)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesr2 <- Pathogens_data %>% filter(Group == "DNA", sample == "Leaves", Experiment_run == "Run3")

View(data.Leavesr2)

#Mix models - experiment run as random variable
model.FWLr2 <- lmer(Fresh_Biomass ~ Visual_dis_severity + (1|Pond), data = data.Leavesr2)

Anova(model.FWLr2) #Significant effect p.value= 9.862e-05 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWLr2, which=1:6) 
res <- resid(model.FWLr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.03515

# Poshoc
emmeans(model.FWLr2, list(pairwise ~ Visual_dis_severity), adjust = "tukey")
#Significant diff 

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Control - High    44.73 11.3  7.13   3.962  0.0127
 #Control - Low      5.34 11.3  7.13   0.473  0.8857
 #High - Low       -39.38 11.4 29.00  -3.439  0.0049

emmeans(model.FWLr2, list(pairwise ~ Visual_dis_severity), adjust = "bonferroni")

# Pairwise comparisons
mSev <- emmeans(model.FWLr2, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low and high-control
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  81.2 8.95 10.8     61.4      101  a    
 #Low                  120.6 8.95 10.8    100.8      140   b   
 #Control              125.9 6.88  4.0    106.8      145   b

```

## Combined boxplot DNA FWR&FWL
```{r}
#Sub-setting only BIOMASS
data.DNA <- Pathogens_data %>% filter(Group == "DNA")

### Dataset imported from Excel 
#To reorder the factor  
data.DNA$Visual_dis_severity <- ordered(data.DNA$Visual_dis_severity, levels = c("Low", "Control", "High"))


# Final plot 
combined.plot.DNA <- ggplot(data.DNA, aes(x = Visual_dis_severity, y = Fresh_Biomass, fill = Visual_dis_severity)) + 
  geom_boxplot() +
  scale_fill_manual(name = "Disease severity", values = c("palegreen1", "#1F78B4", "hotpink1")) +
      labs(x= "\nDisease severity", y = "\nFresh biomass (g)\n", title= "Samples for DNA extraction") +
  facet_grid(~sample + Experiment_run) + theme_classic() +
    theme(legend.position = "none") + 
      scale_y_cut(breaks=41, scales=c(0, 5)) 
#outlier.shape = NA

# Save plot
ggsave("FWR_FWL_byRun.png", combined.plot.DNA, 
  width = 15,
  height = 12,
  units = "cm",
  dpi = 600)

```


# Analysis BIOMASS samples (results DID change - n.s difference whole group)
## Roots fresh weight
```{r}
#Sub-setting only Fresh weight roots
data.Roots <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots")

View(data.Roots)

#Mix models - experiment run as random variable
model.FWR <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run/Pond), data = data.Roots)

model.FWR <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run) + (1|Pond), data = data.Roots)

Anova(model.FWR) #NO Significant effect treatment p.value= 0.2196 / 0.3421

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWR, which=1:6) 
res <- resid(model.FWR)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.6599

# Poshoc
emmeans(model.FWR, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between Control - Inoculated p-value=0.2508

emmeans(model.FWR, list(pairwise ~ Treatment), adjust = "bonferroni")

```

## Roots dry weight
```{r}

#Mix models - experiment run as random variable
model.DWR <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run/Pond), data = data.Roots)

model.DWR <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run) + (1|Pond), data = data.Roots)

Anova(model.DWR) #NO Significant effect treatment p.value= 0.5641 / 0.4417

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWR, which=1:6) 
resD <- resid(model.DWR)

par(mfrow = c(1, 2)) # combine plots
hist(resD, breaks = 40)

# QQ-plot
qqPlot(resD,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resD) #NOT Normal p-value= 0.03759

# Poshoc
emmeans(model.DWR, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.5782

compare_means(Dry_Biomass ~ Treatment, data = data.Roots, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format= and by run2=0.311 - run3=0.018*

```

##Leaves fresh weight
```{r}
#Sub-setting only Fresh weight roots
data.Leaves <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves")

View(data.Leaves)


#Mix models - experiment run as random variable
model.FWL <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run/Pond), data = data.Leaves)

model.FWL <- lmer(Fresh_Biomass ~ Treatment + (1|Experiment_run) + (1|Pond), data = data.Leaves)

Anova(model.FWL) #NO Significant effect treatment p.value= 0.08388 ./0.2247

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWL, which=1:6) 
resL <- resid(model.FWL)

par(mfrow = c(1, 2)) # combine plots
hist(resL, breaks = 40)

# QQ-plot
qqPlot(resL,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resL) #NOT Normal p-value= 0.02895

# Poshoc
emmeans(model.FWL, list(pairwise ~ Treatment), adjust = "tukey")
#Significant diff between Control - Inoculated p-value=0.1179

compare_means(Fresh_Biomass ~ Treatment, data = data.Leaves, method = "t.test", group.by = "Experiment_run")  # Without grouping p.format=0.01* and by run2=0.4429 - run3=0.0042**

compare_means(Fresh_Biomass ~ Treatment, data = data.Leaves, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format= and by run2=0.4429 - run3=0.0042**

```
##Leaves dry weight
```{r}

#Mix models - experiment run as random variable
model.DWL <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run/Pond), data = data.Leaves)

model.DWL <- lmer(Dry_Biomass ~ Treatment + (1|Experiment_run) + (1|Pond), data = data.Leaves)

Anova(model.DWL) #NO Significant effect treatment p.value= 0.1035 / 0.1993

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWL, which=1:6) 
resDL <- resid(model.DWL)

par(mfrow = c(1, 2)) # combine plots
hist(resDL, breaks = 40)

# QQ-plot
qqPlot(resDL,
       id = FALSE # id = FALSE to remove point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(resDL) #NOT Normal p-value= 0.005677


# Poshoc
emmeans(model.DWL, list(pairwise ~ Treatment))
#Significant diff between Control - Inoculated p-value=0.1380

#Non-parametric test
compare_means(Dry_Biomass ~ Treatment, data = data.Leaves, method = "kruskal.test", group.by = "Experiment_run")  # Without grouping p.format=0.022* and by run2=0.547 - run3=0.0056*


# Conduct pairwise comparisons with Wilcoxon rank-sum test
pairwise_result2 <- pairwise.wilcox.test(data.Leaves$Fresh_Biomass, data.Leaves$Treatment, p.adjust.method = "bonferroni")
print(pairwise_result2) #p-value=0.014

```

## Analysis Root length run 2&3 
```{r}
library(readxl)
Pathogens_data2 <- read_excel("Pathogens_RECORD_exps.xlsx",
    sheet = "Pythium_all")
View(Pathogens_data2)
```

```{r}

##### Root length #####
data.RL <- Pathogens_data2 %>% filter(Group == "DNA", Experiment_run != "Run1")

model.RL23 <- lmer(Root_length ~ Visual_dis_severity + (1|Experiment_run/Pond), data = data.RL)

Anova(model.RL23) #Significant effect of Disease severity p-value=0.01645 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.RL23, which=1:6) 
res <- resid(model.RL23)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal   p-value = 0.6147

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.RL23, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  57.0 4.11 2.40     41.9     72.2  a    
 #Control               60.1 3.74 1.64     40.1     80.0  ab   
 #Low                   66.7 4.11 2.40     51.5     81.8   b
 
emmeans(model.RL23, list(pairwise ~ Visual_dis_severity), adjust = "tukey")

#$`pairwise differences of Visual_dis_severity`
 #              estimate   SE   df t.ratio p.value
 #Control - High     3.01 3.92 13.6   0.768  0.7279
 #Control - Low     -6.64 3.92 13.6  -1.695  0.2424
 #High - Low        -9.65 3.42 59.0  -2.820  0.0177 **

```
### Box plot run2&3
```{r}
#To reorder the factor  
data.RL$Visual_dis_severity <- ordered(data.RL$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
ColorsBP <- c("palegreen1", "#1F78B4", "hotpink1")

RL23.plot <- ggboxplot(data.RL, x = "Visual_dis_severity", y = "Root_length",
          fill = "Visual_dis_severity", palette = ColorsBP, outlier.shape = NA) +
      labs(x = "\nDisease severity", y = "Root length (cm)") +
        ylim(25, 100) +
  theme(legend.position = "none") 

# Save plot
ggsave("RL_DNAsamples.png", RL23.plot, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```

# Biomass By RUN
## Roots fresh weight
### Run 1 (run2 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsbr1 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots", Experiment_run == "Run2")

View(data.Rootsbr1)

#Mix models - experiment run as random variable
model.FWRbr1 <- lmer(Fresh_Biomass ~ Treatment + (1|Pond), data = data.Rootsbr1)

Anova(model.FWRbr1) #NO Significant effect treatment p.value= 0.9469

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWRbr1, which=1:6) 
res <- resid(model.FWRbr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.8958

# Poshoc
emmeans(model.FWRbr1, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.9501

emmeans(model.FWRbr1, list(pairwise ~ Treatment), adjust = "bonferroni")

```
### Run 2 (run3 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsbr2 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots", Experiment_run == "Run3")

View(data.Rootsbr2)

#Mix models - experiment run as random variable
model.FWRbr2 <- lmer(Fresh_Biomass ~ Treatment + (1|Pond), data = data.Rootsbr2)

Anova(model.FWRbr2) #Significant effect treatment p.value= 0.03675 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWRbr2, which=1:6) 
res <- resid(model.FWRbr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.3133

# Poshoc
emmeans(model.FWRbr2, list(pairwise ~ Treatment), adjust = "t.test")
#NO Significant diff between Control - Inoculated p-value=0.1050

emmeans(model.FWRbr2, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.1050

```
## Leaves fresh weight
### Run 1 (run2 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesbr1 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves", Experiment_run == "Run2")

View(data.Leavesbr1)

#Mix models - experiment run as random variable
model.FWLbr1 <- lmer(Fresh_Biomass ~ Treatment + (1|Pond), data = data.Leavesbr1)

Anova(model.FWLbr1) #NO Significant effect treatment p.value= 0.4969

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWLbr1, which=1:6) 
res <- resid(model.FWLbr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.0808

# Poshoc (not needed)
emmeans(model.FWLbr1, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.5341

```

### Run 2 (run3 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesbr2 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves", Experiment_run == "Run3")

View(data.Leavesbr2)

#Mix models - experiment run as random variable
model.FWLbr2 <- lmer(Fresh_Biomass ~ Treatment + (1|Pond), data = data.Leavesbr2)

Anova(model.FWLbr2) #NO Significant effect treatment p.value= 0.1252

#Residuals plots
par(mfrow = c(2, 3))
plot(model.FWLbr2, which=1:6) 
res <- resid(model.FWLbr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.1992

# Poshoc (not needed)
emmeans(model.FWLbr2, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.2000

```

## Roots dry weight
### Run 1 (run2 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsbr1 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots", Experiment_run == "Run2")

View(data.Rootsbr1)

#Mix models - experiment run as random variable
model.DWRbr1 <- lmer(Dry_Biomass ~ Treatment + (1|Pond), data = data.Rootsbr1)

Anova(model.DWRbr1) #NO Significant effect treatment p.value= 0.7125

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWRbr1, which=1:6) 
res <- resid(model.DWRbr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.1189

# Poshoc
emmeans(model.DWRbr1, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.7312

```
### Run 2 (run3 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Rootsbr2 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Roots", Experiment_run == "Run3")

View(data.Rootsbr2)

#Mix models - experiment run as random variable
model.DWRbr2 <- lmer(Dry_Biomass ~ Treatment + (1|Pond), data = data.Rootsbr2)

Anova(model.DWRbr2) #Almost Significant effect treatment p.value= 0.0663 .

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWRbr2, which=1:6) 
res <- resid(model.DWRbr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.2796

# Poshoc
emmeans(model.DWRbr2, list(pairwise ~ Treatment), adjust = "t.test")
#NO Significant diff between Control - Inoculated p-value=0.1402


```

## Leaves dry weight
### Run 1 (run2 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesbr1 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves", Experiment_run == "Run2")

View(data.Leavesbr1)

#Mix models - experiment run as random variable
model.DWLbr1 <- lmer(Dry_Biomass ~ Treatment + (1|Pond), data = data.Leavesbr1)

Anova(model.DWLbr1) #NO Significant effect treatment p.value= 0.8613

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWLbr1, which=1:6) 
res <- resid(model.DWLbr1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #NOT Normal p-value=0.02335

# Poshoc (not needed)
emmeans(model.DWLbr1, list(pairwise ~ Treatment), adjust = "tukey")
#NO Significant diff between Control - Inoculated p-value=0.8697

```

### Run 2 (run3 in dataset)
```{r}
#Sub-setting only Fresh weight roots
data.Leavesbr2 <- Pathogens_data %>% filter(Group == "BIOMASS", sample == "Leaves", Experiment_run == "Run3")

View(data.Leavesbr2)

#Mix models - experiment run as random variable
model.DWLbr2 <- lmer(Dry_Biomass ~ Treatment + (1|Pond), data = data.Leavesbr2)

Anova(model.DWLbr2) #ALMOST Significant effect treatment p.value= 0.04576 *

#Residuals plots
par(mfrow = c(2, 3))
plot(model.DWLbr2, which=1:6) 
res <- resid(model.DWLbr2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE # point identification
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal p-value=0.2229

# Poshoc (not needed)
emmeans(model.DWLbr2, list(pairwise ~ Treatment), adjust = "t.test")
#NO Significant diff between Control - Inoculated p-value=0.1164

```

## Combined boxplot Biomass
```{r}
#Sub-setting only BIOMASS
data.Biomass <- Pathogens_data %>% filter(Group == "BIOMASS")

# Final plot Fresh Biomass
combined.plot <- ggplot(data.Biomass, aes(x = sample, y = Fresh_Biomass, fill = Treatment)) + 
  geom_boxplot() +
  facet_grid(~Experiment_run) +
  scale_fill_manual(name = "Pythium inoculation", values = c("#1F78B4", "orchid")) +
  labs(x = "\nSample type", y = "\nFresh biomass (g)\n") +
      theme_bw() +
  scale_y_cut(breaks=37, scales=c(0, 5))


# Save plot
ggsave("Fresh_Biomass_byRun_Pythium.png", combined.plot, 
  width = 12,
  height = 12,
  units = "cm",
  dpi = 600)

```

```{r}

# Final plot Dry Biomass
combined.plotD <- ggplot(data.Biomass, aes(x = sample, y = Dry_Biomass, fill = Treatment)) + 
  geom_boxplot() +
  facet_wrap(~Experiment_run) +
  scale_fill_manual(name = "Pythium inoculation", values = c("#1F78B4", "orchid")) +
  labs(x = "\nSample type", y = "\nDry biomass (g)\n") +
      theme_bw() + 
        scale_y_cut(breaks=3, scales=c(0, 0.5)) 


# Save plot
ggsave("Dry_Biomass_byRun_Pythium.png", combined.plotD, 
  width = 12,
  height = 12,
  units = "cm",
  dpi = 600)

```

## Root length by run 

### Run 1 (run2 in dataset)
```{r}

data.RL <- Pathogens_data %>% filter(Group == "DNA", Experiment_run != "Run1")

data.RL1 <- data.RL %>% filter(sample== "Roots", Experiment_run == "Run2")
View(data.RL1) #36 data observations

model.RL1 <- lmer(Root_length ~ Visual_dis_severity + (1|Pond), data = data.RL1)

Anova(model.RL1) #Significant effect of Disease severity p-value=0.0008297 ***

#Residuals plots
par(mfrow = c(2, 3))
plot(model.RL1, which=1:6) 
res <- resid(model.RL1)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal   p-value = 0.3722

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.RL1, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  53.4 2.09 13.5     48.9     57.9  a    
 #Control               56.6 1.50  4.0     52.5     60.8  a    
 #Low                   64.0 2.09 13.5     59.5     68.5   b 
 
emmeans(model.RL1, list(pairwise ~ Visual_dis_severity), adjust = "tukey")

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Control - High     3.24 2.57  8.41   1.259  0.4529
 #Control - Low     -7.36 2.57  8.41  -2.862  0.0475 *
 #High - Low       -10.60 2.91 29.00  -3.639  0.0029 **

```
### Run 2 (run3 in dataset)
```{r}

#data.RL <- Pathogens_data %>% filter(Group == "DNA", Experiment_run != "Run1")

data.RL2 <- data.RL %>% filter(sample== "Roots", Experiment_run == "Run3")
View(data.RL2) #36 data observations

model.RL2 <- lmer(Root_length ~ Visual_dis_severity + (1|Pond), data = data.RL2)

Anova(model.RL2) #NO Significant effect of Disease severity p-value=0.3711

#Residuals plots
par(mfrow = c(2, 3))
plot(model.RL2, which=1:6) 
res <- resid(model.RL2)

par(mfrow = c(1, 2)) # combine plots
hist(res, breaks = 40)

# QQ-plot
qqPlot(res,
       id = TRUE
      )

# Normality test (normal if p>0.05)
shapiro.test(res) #Normal   p-value = 0.4738

#Perform the test
# Pairwise comparisons
mSev <- emmeans(model.RL2, "Visual_dis_severity")
pairs(mSev) #significant differences between high-low
multcomp::cld(object = mSev, Letters = letters)

#Visual_dis_severity emmean   SE   df lower.CL upper.CL .group
 #High                  60.7 5.95 7.48     46.8     74.6  a    
 #Control               63.5 5.07 4.00     49.4     77.5  a    
 #Low                   69.4 5.95 7.48     55.5     83.3  a  
 
emmeans(model.RL2, list(pairwise ~ Visual_dis_severity), adjust = "tukey")

#$`pairwise differences of Visual_dis_severity`
 #1              estimate   SE    df t.ratio p.value
 #Low - Control      5.92 7.82  5.64   0.757  0.7417
 #Low - High         8.70 6.25 29.00   1.391  0.3587
 #Control - High     2.78 7.82  5.64   0.356  0.9334

```

### Combined Box plot
```{r}
#To reorder the factor  
data.RL$Visual_dis_severity <- ordered(data.RL$Visual_dis_severity, levels = c("Low", "Control", "High"))

# Box plot 
ColorsBP <- c("palegreen1", "#1F78B4", "hotpink1")

RL.plot <- ggboxplot(data.RL, x = "Visual_dis_severity", y = "Root_length",
          fill = "Visual_dis_severity", palette = ColorsBP, outlier.shape = NA) +
      labs(x = "\nDisease severity", y = "Root length (cm)") +
        ylim(25, 100) +
  facet_grid(~Experiment_run) +
  theme(legend.position = "none") 

# Save plot
ggsave("RL_DNAsamples_byRun.png", RL.plot, 
  width = 11,
  height = 12,
  units = "cm",
  dpi = 600)

```


